---
title: "Calibration for GLM Models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calibration for GLM Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
rm(list=ls())
library(RobinCar)
```

## Linear and Joint Calibration for Guaranteed Efficiency Gain and Universality

The dataset that we will use for this is included in the package.

```{r}
data <- RobinCar:::data_sim
data$A <- as.factor(data$A)
```

## Adjustment with Binary Outcomes

Dichotomize the continuous outcome.
```{r}
data$y_bin <- ifelse(data$y > 2, 1, 0)
```

Fit a heterogeneous working model with biased coin randomization,
then perform linear and joint calibration.

```{r}
result <- robincar_glm(
  df=data,
  response_col="y_bin",
  treat_col="A",
  strata_cols=c("z1"),
  covariate_cols=c("x3"),
  car_scheme="biased-coin",
  g_family=binomial(link="logit"),
  covariate_to_include_strata=TRUE,
  g_accuracy=7,
  adj_method="homogeneous"
)
print(result)
```

### Calibration

Use `joint = FALSE` to indicate linear calibration. This fits a model with the predicted potential outcomes as covariates.

```{r}
robincar_calibrate(
  result=result,
  joint=FALSE
)
```

Use `joint = TRUE` to indicate joint calibration. This fits a model with the predicted potential outcomes *and* the joint strata variables.

```{r}
robincar_calibrate(
  result=result,
  joint=TRUE
)
```
