---
title: "Calibration for GLM Models"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
rm(list=ls())
library(RobinCar)
```

## Linear and Joint Calibration for Guaranteed Efficiency Gain and Universality

The dataset that we will use for this is included in the package.

```{r}
data <- RobinCar:::data_sim
data$A <- as.factor(data$A)
 # add some noise as an extra covariate as toy example
data$extra_cov <- rnorm(nrow(data))
```

## Adjustment with Binary Outcomes

Dichotomize the continuous outcome.
```{r}
data$y_bin <- ifelse(data$y > 2, 1, 0)
```

Fit a heterogeneous working model with biased coin randomization,
then perform linear and joint calibration.

```{r}
result <- robincar_glm(
  df=data,
  response_col="y_bin",
  treat_col="A",
  car_strata_cols=c("z1"),
  car_scheme="biased-coin",
  g_family=binomial(link="logit"),
  g_accuracy=7,
  formula="y ~ A + x3 + z1"
)
print(result)
```

### Calibration

Use `joint = FALSE` to indicate linear calibration. This fits a model with the predicted potential outcomes as covariates.

```{r}
robincar_calibrate(
  result=result,
  joint=FALSE
)
```

Use `joint = TRUE` to indicate joint calibration. This fits a model with the predicted potential outcomes *and* the joint strata variables.

```{r}
robincar_calibrate(
  result=result,
  joint=TRUE
)
```

Use `add_x` to include additional covariates in the calibration. This fits a model with the predicted potential outcomes and specified additional covariates (and the joint strata if `joint = TRUE`).

```{r}
robincar_calibrate(
  result=result,
  joint=TRUE,
  add_x=c("extra_cov")
)
```
