---
title: "Continuous and GLM-Type Outcomes"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
rm(list=ls())
library(RobinCar)
```

#### References

Ting Ye, Jun Shao, Yanyao Yi, Qinyuan Zhao (2022+). [Toward Better Practice of Covariate Adjustment in Analyzing Randomized Clinical Trials](https://www.google.com/url?q=https%3A%2F%2Farxiv.org%2Fpdf%2F2009.11828.pdf&sa=D&sntz=1&usg=AOvVaw0OHre3-8OSKzqlaR2yaVrY). *Journal of the American Statistical Association*, in press.

Ting Ye, Marlena Bannick, Yanyao Yi, Jun Shao (2023). [Robust variance estimation for covariate-adjusted unconditional treatment effect in randomized clinical trials with binary outcomes](https://www.tandfonline.com/doi/full/10.1080/24754269.2023.2205802). *Statistical Theory and Related Fields*, 7(2):159-163.

Marlena Bannick, Jun Shao, Jingyi Liu, Yu Du, Yanyao Yi, Ting Ye (2023+). [A General Form of Covariate Adjustment in Randomized Clinical Trials](https://arxiv.org/abs/2306.10213).

## Adjustment with Continuous Outcomes

The dataset that we will use for this is included in the package.

```{r}
data <- RobinCar:::data_sim
data$A <- as.factor(data$A)
```

### ANOVA Model

```{r}
fit.anova <- robincar_linear(df = data, 
                              response_col="y",
                              treat_col="A",
                              covariate_cols=c("x1", "x3"),
                              car_scheme="simple",
                              adj_method="ANOVA")
```

### ANCOVA model with biased coin randomization

```{r}
fit.ancova <- robincar_linear(df = data, 
                               response_col="y",
                               treat_col="A",
                               car_strata_cols=c("z1", "z2"),
                               covariate_cols=c("x1", "x3"),
                               car_scheme="biased-coin",
                               adj_method="ANCOVA")

```

### ANHECOVA model with Pocock-Simon randomization and linear contrast

Include strata variables in the covariates.

```{r}
fit.anhecova <- robincar_linear(df = data, 
                                 response_col="y",
                                 treat_col="A",
                                 car_strata_cols=c("z1", "z2"),
                                 covariate_cols=c("x1", "x3", "z1", "z2"),
                                 car_scheme="pocock-simon",
                                 adj_method="ANHECOVA",
                                 contrast_h="diff")
```

## Adjustment with Binary Outcomes

Dichotomize the continuous outcome.
```{r}
data$y_bin <- ifelse(data$y > 2, 1, 0)
```

### Homogeneous working model with simple randomization

```{r}
glm.homogeneous<-robincar_glm(df = data,
                               response_col="y_bin",
                               treat_col="A",
                               car_strata_cols=c("z1"),
                               formula="y_bin ~ A",
                               car_scheme="pocock-simon",
                               g_family=binomial(link="logit"),
                               g_accuracy=7)
glm.homogeneous
```

### Heterogeneous working model with permuted-block randomization, using one covariate

```{r}
glm.heterogeneous<-robincar_glm(df = data,
                               response_col="y_bin",
                               treat_col="A",
                               car_strata_cols=c("z2"),
                               formula="y_bin ~ A * (x1 + z2)",
                               car_scheme="biased-coin",
                               g_family=binomial(link="logit"),
                               g_accuracy=7)

glm.heterogeneous
```


### Heterogeneous working model with biased-coin randomization, differences contrast

```{r}
glm.heterogeneous <- robincar_glm(df = data,
                                   response_col="y_bin",
                                   treat_col="A",
                                   car_strata_cols=c("z2"),
                                   formula="y_bin ~ A * (x1 + z2)",
                                   car_scheme="biased-coin",
                                   g_family=binomial(link="logit"),
                                   g_accuracy=7,
                                   contrast_h="diff")
glm.heterogeneous
```
