---
title: "RobinCar Family"
author: "Marlena Bannick"
date: "2026-01-19"
output:
  html_document: default
  pdf_document: default
---

This vignette accompanies the Results section of the RobinCar Family paper.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(RobinCar)
library(RobinCar2)
library(dplyr)
library(ranger)
library(SuperLearner)
```
## Case Study 1: ACTG175 Dataset

**Data Manipulation**

We use data from the Juraska et al. (2022) R package, which comes from a
study of nucleoside treatment regimens for individuals with HIV-1 (Hammer et al.
1996) called the AIDS Clinical Trials Group Study 175. This trial had
stratified permuted block randomization.

```{r}
# Data are from the speff2trial package
data <- tibble(speff2trial::ACTG175)

data <- data %>% mutate(
  # Create continuous and binary outcomes
  y_cont = cd420 - cd40,
  y_bin = as.numeric((y_cont / cd40) > 0.5)
) %>% filter(
  # Focus on two treatment arms
  arms %in% c("0", "1")
) %>% mutate(
  # Factor variables for treatment
  # and stratification variable
  arms = factor(arms),
  strat = factor(strat)
)
```

**Linear Adjustment**

The following shows a linear model with treatment-by-covariate interactions, including strata as covariates,
and using the four covariates of weight, hemophilia status, and prior use of non-zidovudine antiretroviral therapy. These specifications ensure that the covariate adjustment has guaranteed efficiency gain (asymptotically).

```{r}
robin_lm(
  y_cont ~ arms * (strat + wtkg + hemo + oprior), 
  treatment = arms ~ pb(strat), 
  data=data)
```

**Generalized Linear Models**

For the binary outcome, we can use a logistic regression model instead.
Rather than a linear contrast, we could instead estimate a log risk ratio. This does not change the marginal mean estimates, but it changes the contrast estimates (and p-values).

```{r}
robin_glm(
  y_bin ~ arms * (strat + wtkg + hemo + oprior),
  treatment = arms ~ pb(strat),
  family = binomial(link = "logit"),
  contrast = "log_risk_ratio",
  data=data
)
```

**Mantel-Haenszel**

RobinCar includes functions for robust Mantel-Haenszel estimation. Here we create a new variable that is the interaction of prior use of non-zidovudine antiretroviral therapy and hemophilia status, and use that as the stratification variable.

By default, the function uses the average treatment effect as the estimand of interest. Specifying the estimand argument as "MH" does not
change the estimate, but it does change the standard error calculation.

The MH function in RobinCar is only valid for simple randomization.

```{r}
data <- data %>% mutate(
  X=interaction(oprior, hemo)
)

robincar_mh(
  data, "arms", "y_bin",
  strata_cols=c("oprior", "hemo")
)
```

**Survival Analysis**

For a survival outcome with right-censoring, we can use the RobinCar2 function `robin_surv`
to do a covariate-adjusted, stratified log-rank test.

```{r}
surv <- robin_surv(
  Surv(days, cens) ~ wtkg + oprior + hemo + strat,
  treatment = arms ~ pb(strat),
  data = data
)
surv
```

You can use the `table` function to see the number of events and number at-risk

```{r}
table(surv)
```

## Case Study 2: RCT of Indomethacin for Post-ERCP Pancreatitis

**Data Manipulation**

This case study uses a different dataset, `medicaldata::indo_rct`, which was
a randomized trial of rectal indomethacin to prevent post-ERCP pancreatitis
(Elmunzer, Higgins, et al. 2012). This study used simple randomization
stratified by study site (this is not the same as stratified permuted block
randomization in the first example).

```{r}
require(medicaldata)

data2 <- medicaldata::indo_rct %>%
  mutate(
    outcome2=as.numeric(outcome)-1
  ) %>%
  select(
    -c(outcome, status, type, bleed)
  )
```

First fit a really simple logistic regression working model only adjusting 
for site, and a risk difference as our contrast of interest.

```{r}
robin_glm(
  outcome2 ~ rx * site, 
  treatment = rx ~ sr(1), 
  data=data2,
  family=binomial(link="logit"),
  contrast = "difference"
)
```

There are 27 baseline covariates included in this dataset and 602 patients.
If we include all of the covariates in a logistic regression working model,
the model fails to converge and returns NA.

```{r}
# Create formula with all baseline covariates included in dataset
covariates <- setdiff(colnames(data2), c("id", "outcome2", "rx"))
formula <- as.formula(
  sprintf("outcome2 ~ rx * (%s)", paste0(covariates, collapse=" + "))
)
```

```{r}
robin_glm(
  as.formula(formula), 
  treatment = rx ~ sr(1), 
  data=data2,
  family=binomial(link="logit"),
  contrast = "difference"
)
```

If you do want to include all of the baseline covariates because you aren't
sure which ones are the most prognostic, this is a use-case for the
SuperLearner approach. We'll test an ensemble
of random forest (SL.ranger), elastic net regression (Sl.glmnet),
and XGboost (SL.xgboost). SuperLearner
will figure out the optimal combination of these approaches.
See [this link](https://cran.r-project.org/web/packages/SuperLearner/vignettes/Guide-to-SuperLearner.html) for documentation on the SuperLearner.

RobinCar includes SuperLearner working models
with cross-fitting to ensure we aren't over-fitting to the data.
We will use 5 folds for cross-fitting. Since cross-fitting is a random
process, we'll also set a seed for reproducibility. 

```{r}
set.seed(1989)

sl <- robincar_SL(
  df=data2,
  response_col="outcome2",
  treat_col="rx",
  covariate_cols=covariates,
  SL_libraries=c("SL.xgboost", "SL.glmnet", "SL.ranger"),
  car_scheme="simple",
  k_split=5
)

# Make a linear contrast
contrast <- robincar_contrast(sl, contrast_h="diff")
contrast
```
